---
title: "Chapter 11 - Ice Cream Ratings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gsheet)
library(tidyverse)

data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1QH5XDEUEmBGp2ZzO7OQS0T1vSELIjFAi2gTmnqokoMg/edit#gid=0"))
```

Commercially produced ice cream is made from a mixture of
ingredients: a minimum of 10% milk fat, 9—12% milk solids,
12—16% sweetener, 0.2—0.5% stabilizers and emulsifiers, and
55—64% water. Air is incorporated with the above ingredients
during the mixing process. The finest ice creams have between 3%
and 15% air.

A food scientist is investigating how varying the amounts of the
above ingredients impacts the sensory rating of the final product.
The scientist decides to use three levels of milk fat: 10%, 12%,
and 15%; three amounts of air: 5%, 10%, 15%; and two levels of
sweeteners: 12%, 16%.

Estimate the HWF model with a three-way interaction.

#### Construct the Model with the Three-Way Interaction

```{r}
data$Sweetner <- str_remove(data$Sweetner, "[%]")
  data$Sweetner <- as.numeric(data$Sweetner)

data$MilkFat <- str_remove(data$MilkFat, "[%]")
  data$MilkFat <- as.numeric(data$MilkFat)
  
data$Air <- str_remove(data$Air, "[%]")
  data$Air <- as.numeric(data$Air)

m1 <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air + Sweetner:MilkFat:Air, data=data)

summary(m1)
```

#### Testing the Three-Way Interaction Between Sweetner, Milk Fat, and Air

**Hypotheses**

&emsp;&emsp; $H_0:$ there is not an interaction between sweetner, milk fat, and air <br>
&emsp;&emsp; $H_1:$ there is an interaction between sweetner, milk fat, and air

**Test Statistic**

&emsp;&emsp; $t_0$ = -0.545

***p*-value**

&emsp;&emsp; $p$ = 0.588

**Rejection Region**

&emsp;&emsp; Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

&emsp;&emsp; Fail to reject $H_0$. There is not sufficient evidence to suggest that there is an interaction between sweetner, milk fat, and air.

#### Remove the Three-Way Interaction

```{r}
m2 <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air, data=data)

summary(m2)
```


#### Test Significance of All Interaction Terms

```{r}
full <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air, data=data)

reduced <- lm(Ratings ~ Sweetner + MilkFat + Air, data=data)

anova(reduced, full)
```

**Hypotheses**

&emsp;&emsp; $H_0:$ there are no significant two-way interactions <br>
&emsp;&emsp; $H_1:$ at least one of the two-way interactions is significant

**Test Statistic**

&emsp;&emsp; $F_0$ = 3.108

***p*-value**

&emsp;&emsp; $p$ = 0.035

**Rejection Region**

&emsp;&emsp; Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

&emsp;&emsp; Reject $H_0$. There is sufficient evidence to suggest that there is at least one significant two-way interaction.

#### Examining All Two-Way Interactions

```{r}
summary(m2)
```
Only the interaction between milk fat and air is significant (*p* = 0.011). The interactions between sweetner and milk fat and, separately, sweetner and air are not significant (*p* = 0.840 and *p* = 0.131, respectively). 

Can we remove the non-significant interactions at the same time?

```{r}
full <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air, data=data)

reduced <- lm(Ratings ~ Sweetner + MilkFat + Air + MilkFat:Air, data=data)

anova(reduced, full)
```

Yes, we can remove both at the same time (*p* = 0.311). For parsimony, we will go ahead and remove those two interaction terms.

#### Model with Interaction Between Milk Fat and Air

```{r}
m3 <- lm(Ratings ~ Sweetner + MilkFat + Air + MilkFat:Air, data=data)

summary(m3)
```
Because the interaction between milk fat and air is in the model, we are required to keep the main effects for milk fat and air in the model. However, we see that sweetner is not significant (*p* = 0.147) and it is not involved in an interaction term... so it can go.

#### Final Model

```{r}
m4 <- lm(Ratings ~ MilkFat + Air + MilkFat:Air, data=data)

summary(m4)
```

#### Visualization

We now will construct a visualization. First, we must decide what goes where on the graph. Let's put milk fat on the *x* axis and create lines with air. Let's remind ourselves of the levels of air:

```{r}
data %>% count(Air)
```

We will use levels 5, 10, and 15 in our predictions. Note that if this were truly a continuous variable, we would want to look at the 25th percentile, the median, and the 75th percentile. 

```{r}
c <- coefficients(m4)

data <- data %>% mutate(
          air5 = c[1] + c[2]*MilkFat + c[3]*5 + c[4]*5*MilkFat,
          air10 = c[1] + c[2]*MilkFat + c[3]*10 + c[4]*10*MilkFat,
          air15 = c[1] + c[2]*MilkFat + c[3]*15 + c[4]*15*MilkFat
        )
```

We also will color our observations on the graph by the level of sweetner. We first need to turn it back into a factor variable,

```{r}
data$Sweetner <- as.factor(data$Sweetner)
```

Now we are ready to graph,

```{r}
ggplot(data, aes(x=MilkFat, y=Ratings, color = Sweetner)) +
        geom_point() +
        geom_line(aes(y = air5), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 35.5, label = "5% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = air10), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 32.3, label = "10% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = air15), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 29.25, label = "15% Air"), color="black", show.legend = FALSE) +
        xlab("Milk Fat (%)") +
        ylab("Rating of Ice Cream") +
        theme_minimal() 
```

We can see that as milk fat increases, ratings increase with 5% air, slightly increase with 10% air, and decrease with 15% air.

We can also examine this the other way, with air on the *x*-axis and milk fat defining the lines. Let's look at the levels of milk fat.

```{r}
data %>% count(MilkFat)
```

We will use levels 10, 12, and 15 in our predictions. Like we noted for air, if this were truly a continuous variable, we would want to look at the 25th percentile, the median, and the 75th percentile.

```{r}
data <- data %>% mutate(
          mf10 = c[1] + c[2]*10 + c[3]*Air + c[4]*Air*10,
          mf12 = c[1] + c[2]*12 + c[3]*Air + c[4]*Air*12,
          mf15 = c[1] + c[2]*15 + c[3]*Air + c[4]*Air*15
        )
```

Now we are ready to graph,

```{r}
ggplot(data, aes(x=Air, y=Ratings, color = Sweetner)) +
        geom_point() +
        geom_line(aes(y = mf10), color = "black", linetype = "solid") +
        geom_text(aes(x = 16.2, y = 33, label = "10% Milk Fat"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = mf12), color = "black", linetype = "solid") +
        geom_text(aes(x = 16.2, y = 31.5, label = "12% Milk Fat"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = mf15), color = "black", linetype = "solid") +
        geom_text(aes(x = 16.2, y = 29.25, label = "15% Milk Fat"), color="black", show.legend = FALSE) +
        xlab("Air (%)") +
        ylab("Rating of Ice Cream") +
        xlim(4,17) +
        theme_minimal() 
```

We can see that as air increases, ratings increase with 10% milk fat, stay the same with 12% milk fat, and decrease with 15% milk fat.