---
title: "Chapter 12 - Ice Cream Ratings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gsheet)
library(fastDummies)
library(tidyverse)
library(ggpubr)

data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1QH5XDEUEmBGp2ZzO7OQS0T1vSELIjFAi2gTmnqokoMg/edit#gid=0"))
```

Commercially produced ice cream is made from a mixture of ingredients: a minimum of 10% milk fat, 9—12% milk solids, 12—16% sweetener, 0.2—0.5% stabilizers and emulsifiers, and 55—64% water. Air is incorporated with the above ingredients during the mixing process. The finest ice creams have between 3% and 15% air.

A food scientist is investigating how varying the amounts of the above ingredients impacts the sensory rating of the final product. The scientist decides to use three levels of milk fat: 10%, 12%, and 15%; three amounts of air: 5%, 10%, 15%; and two levels of sweeteners: 12%, 16%.

Estimate the HWF model with a three-way interaction.

#### Construct the Model with the Three-Way Interaction

```{r}
m1 <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air + Sweetner:MilkFat:Air, data=data)

summary(m1)
```

We can see that by allowing the predictors to be categorical predictors, it complicates / adds more predictors to our model. To test the three-way interaction in the overall test, we will turn to the multiple partial *F*.

```{r}
full <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air + Sweetner:MilkFat:Air, data=data)
reduced <- lm(Ratings ~ Sweetner + MilkFat + Air + Sweetner:MilkFat + Sweetner:Air + MilkFat:Air, data=data)

anova(reduced, full)
```

#### Testing the Three-Way Interaction Between Sweetner, Milk Fat, and Air

**Hypotheses**

&emsp;&emsp; $H_0:$ there is not an interaction between sweetner, milk fat, and air <br>
&emsp;&emsp; $H_1:$ there is an interaction between sweetner, milk fat, and air

**Test Statistic**

&emsp;&emsp; $F_0$ = 3.523

***p*-value**

&emsp;&emsp; $p$ = 0.016

**Rejection Region**

&emsp;&emsp; Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

&emsp;&emsp; Reject $H_0$. There is sufficient evidence to suggest that there is a three-way interaction between sweetner, milk fat, and air.

### Models

Because the three-way interaction exists, let's make it a little simpler -- let's pick one variable to plug in for and simplify our models. 

Let's hold sweetner constant -- there are two levels. The models are then,
\begin{align*}
\text{S=12:} & \ \hat{y} = 35.67 + 1.33 \text{MF}_{\text{12}} - 1.67 \text{MF}_{\text{15}} - 9.33 \text{A}_{\text{15}} - 11.67 \text{A}_{\text{5}} + 1.67 \text{MF}_{\text{12}} \text{A}_{\text{5}} + 8.33 \text{MF}_{\text{15}} \text{A}_{\text{5}} + 7.00 \text{MF}_{\text{12}} \text{A}_{\text{15}} + 1.33 \text{MF}_{\text{15}} \text{A}_{\text{15}} \\
\text{S=16:} & \ \hat{y} = 37.00 - 1.00 \text{MF}_{\text{12}} - 3.33 \text{MF}_{\text{15}} - 12.00 \text{A}_{\text{5}} - 10.33 \text{A}_{\text{15}} + 12.33 \text{MF}_{\text{12}} \text{A}_{\text{5}} + 14.67 \text{MF}_{\text{15}} \text{A}_{\text{5}} +  10.00 \text{MF}_{\text{12}} \text{A}_{\text{15}} + 2.67 \text{MF}_{\text{15}} \text{A}_{\text{15}}
\end{align*}

#### Visualization

From here, we can go to visualizations. For brevity, let us focus only on milk fat on the *x* axis and lines defined by air. Note that we will have one graph for sweetner of 12% and another graph for sweetner of 16%. Let's remind ourselves of the levels of air:

```{r}
data %>% count(Air)
```

We will use levels 5, 10, and 15 in our predictions. Like we noted in the last chapter, if this were truly a continuous variable, we would want to look at the 25th percentile, the median, and the 75th percentile. 

We will start with 12% sweetner.

```{r}
data <- dummy_cols(data, select_columns = c("Air", "MilkFat", "Sweetner"))

#35.67 + 1.33 * `MilkFat_12%` - 1.67 * `MilkFat_15%` - 9.33 * `Air_15%` - 11.67 * `Air_5%` + 1.67 * `MilkFat_12%` *`Air_5%` + 8.33 * `MilkFat_15%` * `Air_5%` + 7.00 * `MilkFat_12%` * `Air_15%` + 1.33 * `MilkFat_15%` * `Air_15%`

data <- data %>% mutate(
          s12air5 = 35.67 + 1.33 * `MilkFat_12%` - 1.67 * `MilkFat_15%` - 11.67 * 1 + 1.67 * `MilkFat_12%` *1 + 8.33 * `MilkFat_15%` * 1,
          s12air10 = 35.67 + 1.33 * `MilkFat_12%` - 1.67 * `MilkFat_15%`,
          s12air15 = 35.67 + 1.33 * `MilkFat_12%` - 1.67 * `MilkFat_15%` - 9.33 *1 + 7.00 * `MilkFat_12%` * 1 + 1.33 * `MilkFat_15%` * 1
        )
```

We also want to turn milk fat into a continuous variable for graphing purposes.

```{r}
data$MilkFat <- str_remove(data$MilkFat, "[%]")
  data$MilkFat <- as.numeric(data$MilkFat)
```

Now we are ready to graph,

```{r}
p1 <- ggplot(data, aes(x=MilkFat, y=Ratings)) +
        geom_point() +
        geom_line(aes(y = s12air5), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 30.5, label = "5% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = s12air10), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 34, label = "10% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = s12air15), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 26, label = "15% Air"), color="black", show.legend = FALSE) +
        xlab("Milk Fat (%)") +
        ylab("Rating of Ice Cream") +
        ggtitle("12% Sweetner") +
        theme_minimal() 

p1
```

We can see that as milk fat increases, ratings consistently increase with 5% air, but with both 10% air and 15% air, the ratings increase between 10% and 12% milk fat but decrease between 12% and 14% milk fat.

Now, let's look at the graph for 16% sweetner. First, our predicted values.

```{r}
#37.00 - 1.00* `MilkFat_12%` - 3.33 *`MilkFat_15%` - 12.00 *`Air_5%` - 10.33 *`Air_15%` + 12.33 *`MilkFat_12%`* `Air_5%` + 14.67* `MilkFat_15%` *`Air_5%` +  10.00 *`MilkFat_12%`* `Air_15%` + 2.67* `MilkFat_15%`*`Air_15%`

data <- data %>% mutate(
          s16air5 = 37.00 - 1.00* `MilkFat_12%` - 3.33 *`MilkFat_15%` - 12.00 *1  + 12.33 *`MilkFat_12%`* 1 + 14.67* `MilkFat_15%` *1,
          s16air10 = 37.00 - 1.00* `MilkFat_12%` - 3.33 *`MilkFat_15%`,
          s16air15 = 37.00 - 1.00* `MilkFat_12%` - 3.33 *`MilkFat_15%` - 10.33 *1 +  10.00 *`MilkFat_12%`* 1 + 2.67* `MilkFat_15%`*1
        )
```

Then, our graph.

```{r}
p2 <- ggplot(data, aes(x=MilkFat, y=Ratings)) +
        geom_point() +
        geom_line(aes(y = s16air5), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 36.25, label = "5% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = s16air10), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 33.5, label = "10% Air"), color="black", show.legend = FALSE) + 
        geom_line(aes(y = s16air15), color = "black", linetype = "solid") +
        geom_text(aes(x = 15.3, y = 26.2, label = "15% Air"), color="black", show.legend = FALSE) +
        xlab("Milk Fat (%)") +
        ylab("Rating of Ice Cream") +
        ggtitle("16% Sweetner") +
        theme_minimal() 

p2
```

For 5% air, ratings increase between 10% and 12% milk fat and level off between 12% and 15% milk fat. For 10% air, ratings steadily decrease as milk fat increases. For 15% air, ratings increase between 10% and 12% milk fat and decrease between 12 and 15% milk fat.


Let's look at these graphs side by side:

```{r}
ggarrange(p1, p2, ncol=2)
```

By putting the graphs side by side, we can easily see the differences between the expected values. 

Note that these were "automatically" on the same scale for the *y*-axis ... in the event that the scales were not the same, we should go back and force the graphs to have the same scale. Without the graphs being on the same scale, we cannot easily compare the two models/graphs.


## How to Create Dummy Variables and Specify Reference Groups

Now we will learn how to create our own dummy variables for variables that are not given to us in 0/1 format.
```{r}
# create dummy vars
data <- dummy_cols(data, select_columns = c("Sweetner", "MilkFat", "Air"))
# this creates dummy variables for all of the variables specified
# it will create a column for each level of the variable
# e.g., sweetner will have sweetner_12% and sweetner_16%

# look at variable names
colnames(data)
```

The dummy_cols() function (from the fastDummies package) creates the dummy variables automatically for us. It creates literally one variable for each level of the variable. For example, sweetner has two levels: 12% and 16%. There are now two additional variables: Sweetner_12% and Sweetner_16%. Note that R does not like special characters in the variable names, so for us to be able to appropriately call the variable, we must put back ticks (`) around the variable name.

To specify one of the groups as the reference group, we exclude that variable from the model. For example, if we specify the lowest level of each of the factors as the reference groups, we will exclude Sweetner_12%, MilkFat_10%, and Air_5%. That means that we will include the others: Sweetner_16%, MilkFat_12%, MilkFat_15%, Air_10%, and Air_15%.

```{r}
m1 <- lm(Ratings ~ `Sweetner_16%` + `MilkFat_12%` + `MilkFat_15%` + `Air_10%` + `Air_15%`, data=data)
summary(m1)
```

When specifying interaction terms, we must specify pairwise terms. Suppose that we are interested in the interaction between sweetner and air. We will include all pairs: Sweetner_16%:Air_10% and Sweetner_16%:Air_15%. Note that as the number of levels in a categorical predictor increases, the number of pairs that need to be specified also increases. 
```{r}
m1 <- lm(Ratings ~ `Sweetner_16%` + `MilkFat_12%` + `MilkFat_15%` + `Air_10%` + `Air_15%`
         + `Sweetner_16%`:`Air_10%` + `Sweetner_16%`:`Air_15%`, data=data)
summary(m1)
```

Finally, if we want to test the global significance of the interaction (i.e., the generic "is this interaction significant?"), we must use a multiple partial F test. The full model will contain all predictors, including interaction terms. The reduced model will have the interaction terms of interest removed. Note that if you need to test multiple interactions (e.g., the interaction between sweetner and air and, separately, the interaction between air and milk fat), you will constuct two tests (i.e., two "reduced" models).
```{r}
full <- lm(Ratings ~ `Sweetner_16%` + `MilkFat_12%` + `MilkFat_15%` + `Air_10%` + `Air_15%`
         + `Sweetner_16%`:`Air_10%` + `Sweetner_16%`:`Air_15%`, data=data)
reduced <- lm(Ratings ~ `Sweetner_16%` + `MilkFat_12%` + `MilkFat_15%` + `Air_10%` + `Air_15%`, data=data)
anova(reduced, full)
```

Once we confirm that an interaction is significant at the global level, we can do posthoc testing to determine if specific comparisons are significant (i.e., the individual *t*-test from the summary() function).
```{r}
summary(m1)
```