---
title: 'STA4234: Homework Starter'
output: 
  html_document:
      toc: TRUE
      toc_depth: 4
      toc_float: TRUE
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
```

## Data from 14.22 and 14.23

In an analysis of daily soil evaporation (EVAP), Freund (1979) identified the following predictor variables:
    MAXAT: Maximum daily air temperature 
    MINAT: Minimum daily air temperature
    AVAT: Integrated area under the daily air temperature curve (i.e., a measure of average air temperature)
    MAXST: Maximum daily soil temperature 
    MINST: Minimum daily soil temperature
    AVST: Integrated area under the daily soil temperature curve 
    MAXH: Maximum daily relative humidity
    MINH: Minimum daily relative humidity 
    AVH: Integrated area under the daily relative humidity curve
    WIND: Total wind, measured in miles per day

```{r}
data <- read_sheet("https://docs.google.com/spreadsheets/d/1GZeDx6LZNML2MURS6Nl2ap3V9WQQRCFX-GPkZwHzEqo/edit?usp=sharing")
```




