---
title: 'STA4234: Chapter 8 Examples'
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
      toc: TRUE
      toc_depth: 4
      toc_float: TRUE
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

# info about multiple regression in R here: https://www.statmethods.net/stats/regression.html

```

## Example 1

The following data are weight ($Y$), height ($X_1$), and age ($X_2$) for a sample of 12 children. 

```{r}
wgt <- c(64, 71, 53, 67, 55, 58, 77, 57, 56, 51, 76, 68)
hgt <- c(57, 59, 49, 62, 51, 50, 55, 48, 42, 42, 61, 57)
age <- c(8, 10, 6, 11, 8, 7, 10, 9, 10, 6, 12, 9)

one <- tibble(wgt, hgt, age)

# create age^2 variable
one$age2 <- one$age^2
```

### Model 1 (wgt ~ hgt + age + age2)

```{r}
one_m1 <- lm(wgt ~ hgt + age + age2, data=one)
one_m1_coef <- coefficients(one_m1)
```

#### Model

The resulting regression model is 
\[ \hat{Y} = `r round(one_m1_coef[[1]], digits = 2)` + `r round(one_m1_coef[[2]], digits=2)`X_1 + `r round(one_m1_coef[[3]], digits=2)`X_2 `r round(one_m1_coef[[4]], digits=2)`X_2^2 \]

#### Graph holding height constant

```{r}
# add variables containing y hat
one$pred50_m1 <- one_m1_coef[[1]] + one_m1_coef[[2]]*50 + one_m1_coef[[3]]*one$age + one_m1_coef[[4]]*one$age2
one$pred53_m1 <- one_m1_coef[[1]] + one_m1_coef[[2]]*53 + one_m1_coef[[3]]*one$age + one_m1_coef[[4]]*one$age2
one$pred55_m1 <- one_m1_coef[[1]] + one_m1_coef[[2]]*55 + one_m1_coef[[3]]*one$age + one_m1_coef[[4]]*one$age2

p <- ggplot(one, aes(x=age, y=wgt)) +
  geom_point() +
  geom_line(aes(y = pred50_m1), color = "black", linetype = "dashed") +
  geom_line(aes(y = pred53_m1), color = "black", linetype = "dashed") +
  geom_line(aes(y = pred55_m1), color = "black", linetype = "dashed") +
  geom_text(aes(x = 12.5, y = 70.7, label = "height = 55")) + 
  geom_text(aes(x = 12.5, y = 69.25, label = "height = 53")) +
  geom_text(aes(x = 12.5, y = 67, label = "height = 50")) +
  xlab("Age") + xlim(6, 13) +
  ylab("Weight") +
  theme_bw()

p
```

#### ANOVA table

```{r}
one_anova_m1 <- anova(one_m1)
```

The corresponding ANOVA table is as follows

`r pander(one_anova_m1, style='rmarkdown')`

#### $R^2$ and $R^2_{\mbox{adj}}$

```{r}
one_summary_m1 <- summary(one_m1)
```

$R^2 = `r round(one_summary_m1$r.squared, digits = 4)`$ and $R^2_{\mbox{adj}} = `r round(one_summary_m1$adj.r.squared, digits = 4)`$

### Model 2 (wgt ~ hgt + age)

```{r}
one_m2 <- lm(wgt ~ hgt + age, data=one)
one_m2_coef <- coefficients(one_m2)

one_m2_coef
```

#### Model 

The resulting regression model is 
\[ \hat{Y} = `r round(one_m2_coef[[1]], digits = 2)` + `r round(one_m2_coef[[2]], digits=2)`X_1 + `r round(one_m2_coef[[3]], digits=2)`X_2  \]

#### Graph holding height constant

```{r}
# add variables containing y hat
one$pred50_m2 <- one_m2_coef[[1]] + one_m2_coef[[2]]*50 + one_m2_coef[[3]]*one$age 
one$pred53_m2 <- one_m2_coef[[1]] + one_m2_coef[[2]]*53 + one_m2_coef[[3]]*one$age 
one$pred55_m2 <- one_m2_coef[[1]] + one_m2_coef[[2]]*55 + one_m2_coef[[3]]*one$age 

p <- ggplot(one, aes(x=age, y=wgt)) +
  geom_point() +
  geom_line(aes(y = pred50_m2), color = "black", linetype = "dashed") +
  geom_line(aes(y = pred53_m2), color = "black", linetype = "dashed") +
  geom_line(aes(y = pred55_m2), color = "black", linetype = "dashed") +
  geom_text(aes(x = 12.5, y = 71, label = "height = 55")) + 
  geom_text(aes(x = 12.5, y = 69.5, label = "height = 53")) +
  geom_text(aes(x = 12.5, y = 67.25, label = "height = 50")) +
  xlab("Age") + xlim(6, 13) +
  ylab("Weight") +
  theme_bw()

p
```

#### ANOVA table

```{r}
one_anova_m2 <- anova(one_m2)
```

The corresponding ANOVA table is as follows

`r pander(one_anova_m2, style='rmarkdown')`

#### $R^2$ and $R^2_{\mbox{adj}}$

```{r}
one_summary_m2 <- summary(one_m2)
```

$R^2 = `r round(one_summary_m2$r.squared, digits = 4)`$ and $R^2_{\mbox{adj}} = `r round(one_summary_m2$adj.r.squared, digits = 4)`$






