---
title: 'STA4234: Chapter 7 Examples'
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
      toc: TRUE
      toc_depth: 4
      toc_float: TRUE
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

# info about multiple regression in R here: https://www.statmethods.net/stats/regression.html

```

## Example 1

The following data are systolic blood pressure (SBP) and age for a sample of 30 individuals.

```{r}

sbp <- c(144, 220, 138, 145, 162, 142, 170, 124, 158, 154, 162, 150, 140, 110, 128, 
         130, 135, 114, 116, 124, 136, 142, 120, 120, 160, 158, 144, 130, 125, 175)

age<- c(39, 47, 45, 47, 65, 46, 67, 42, 67, 56, 64, 56, 59, 34, 42, 
        48, 45, 17, 20, 19, 36, 50, 39, 21, 44, 53, 63, 29, 25, 69)

one <- tibble(sbp, age)
```

### ANOVA table 

```{r}
one_model <- lm(sbp ~ age, data=one)
one_anova <- anova(one_model)
```

The corresponding ANOVA table is as follows

`r pander(one_anova, style='rmarkdown')`

### Is the regression line significant?

**Hypotheses**

&emsp;&emsp; $H_0: \ \beta_1 = 0$ <br>
&emsp;&emsp; $H_1: \ \beta_1 \ne 0$

**Test Statistic**

&emsp;&emsp; $F_0 = `r round(one_anova$"F value"[1], digits = 2)`$.

***p*-value**

&emsp;&emsp; $`r p.value.string(one_anova$"Pr(>F)"[1])`$.

**Rejection Region**

&emsp;&emsp; Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

&emsp;&emsp; Reject $H_0$. There is sufficient evidence to suggest that the regression line is significant.


