---
title: 'STA4234: Chapter 14 Examples'
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(pander)
library(car)
library(lindia)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

almost_sas <- function(aov.results){
  par(mfrow=c(2,2))
  plot(aov.results, which=1)
  plot(aov.results, which=2)
  aov_residuals <- residuals(aov.results)
  plot(density(aov_residuals))
  hist(aov_residuals)
}
```

## HuMoSim Example

Let's explore the concepts discussed today. 

```{r}
one <- read_csv("/Volumes/GoogleDrive/My Drive/STA4234/R Markdown/driver.csv")

# create ID variable
one <- mutate(one, obs = row_number())
```

### Correlation

Let's look at the correlation between the variables. Are there any correlations that are abnormally high and we should be worried about?

```{r}
pander(cor(one), style='rmarkdown')
```

### Model 1
Let's construct a basic model with only main effects.

```{r}
m1 <- lm(weight	~ age + htshoes +	ht + seated + arm + thigh + leg, data=one)
pander(summary(m1), style='rmarkdown')
```

#### VIF
Now, let's check the VIF of that model:

```{r}
pander(vif(m1), style='rmarkdown')
```

Should we adjust our model?

### Model 2

```{r}
# Edit the following model as necessary:
#m2 <- lm(weight	~ age + htshoes +	ht + seated + arm + thigh + leg, data=one)
#pander(summary(m2, style='rmarkdown'))
#pander(vif(m2), style='rmarkdown')
```

Should we adjust our model?

### Model 3

```{r}
# Edit the following model as necessary:
#m3 <- lm(weight	~ age + htshoes +	ht + seated + arm + thigh + leg, data=one)
#pander(summary(m3, style='rmarkdown'))
#pander(vif(m3), style='rmarkdown')
```

Should we adjust our model?

#### R Student

Let's check the studentized residuals (note: in R they are called standardized residuals; R uses "studentized" to describe another type of residual, the jackknife residual).

```{r}
pander(rstandard(m3), style='rmarkdown')
```

Should we investigate any points?

#### Cook's D

Let's check influence/leverage using Cook's distance. 

```{r}
gg_cooksd(m3) + theme_bw()
```

Are there any points we should investigate? 

#### Model Assumptions

Let's check the model assumptions. Remember, we assume that the residuals are normally distributed with mean 0 and a common variance. We should use the scatterplot (top left) to assess variance, the Q-Q plot (top right) to assess normality, and the density/histogram (bottom row) to further assess normality.

```{r}
almost_sas(m3)
```

What are your thoughts? Are the assumptions met (roughly)?

### Model 4 

Reconstruct M3 without the points identified as issues.


```{r}
# Exclude the potential issue datapoints
#one_m4 <- filter(one, obs != 1 & obs != 2 & ...)
#m4 <- lm(weight	~ age + seated + arm + thigh + leg, data=one_m4)
#pander(summary(m3), style='rmarkdown')
#pander(summary(m4), style='rmarkdown')
```

How different are the results?

#### Diagnostics

Remember: we only check diagnostics once. If we continue to re-evaluate, we will continue to find issue points (and will whittle down our dataset such that it is no longer generalizable to the initial population of interest).

#### Model Assumptions

Let's check the model assumptions. 

```{r}
almost_sas(m4)
```

What are your thoughts? Are the assumptions met (roughly)?

