---
title: 'STA4234: Chapter 5 Examples'
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
      toc: TRUE
      toc_depth: 4
      toc_float: TRUE
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(pander)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

# info about multiple regression in R here: https://www.statmethods.net/stats/regression.html

```

## Example 1

The following data are systolic blood pressure (SBP) and age for a sample of 30 individuals. We will construct the scatter plot and the regression line of best fit.

```{r}

sbp <- c(144, 220, 138, 145, 162, 142, 170, 124, 158, 154, 162, 150, 140, 110, 128, 
         130, 135, 114, 116, 124, 136, 142, 120, 120, 160, 158, 144, 130, 125, 175)

age<- c(39, 47, 45, 47, 65, 46, 67, 42, 67, 56, 64, 56, 59, 34, 42, 
        48, 45, 17, 20, 19, 36, 50, 39, 21, 44, 53, 63, 29, 25, 69)

one <- tibble(sbp, age)
```

### Summary statistics
```{r}
one.summary <- one %>%
               summarise(mean.sbp = mean(sbp), mean.age = mean(age), sd.sbp = sd(sbp), sd.age = sd(age))
```

The mean (standard deviation) SBP is `r round(one.summary$mean.sbp[1], digits = 2)` (`r round(one.summary$sd.sbp[1], digits = 2)`) mmHg.

The mean (standard deviation) age is `r round(one.summary$mean.age[1], digits = 2)` (`r round(one.summary$sd.age[1], digits = 2)`) years.

### Scatterplot

```{r}
p <- ggplot(one, aes(x=age, y=sbp)) + 
        geom_point() + 
        theme_bw()

p
```

### Regression model

```{r}
one_model <- lm(sbp ~ age, data=one)
one_coef <- coefficients(one_model)
```

The resulting regression model is 
\[ \hat{y} = `r round(one_coef[[1]], digits = 2)` + `r round(one_coef[[2]], digits=2)`x \]

For a 1 year increase in age, we expect SBP to increase by `r round(one_coef[[2]], digits=2)` mmHg. 

When age is 0, we expect SBP to be `r round(one_coef[[1]], digits = 2)` mmHg.

### Estimation of $s^2_{Y|X}$

```{r}
one_anova <- anova(one_model)
```

The corresponding ANOVA table is as follows

`r pander(one_anova, style='rmarkdown')`

Thus, MSE = `r round(one_anova$"Mean Sq"[2], digits = 4)`.

### Is $X$ a significant predictor of $Y$?

```{r}
one_summary <- summary(one_model)
one_t <- as_tibble(one_summary[[4]])
```

**Hypotheses**

&emsp;&emsp; $H_0: \ \beta_1 = 0$ <br>
&emsp;&emsp; $H_1: \ \beta_1 \ne 0$

**Test Statistic**

&emsp;&emsp; $t_0 = `r round(one_t$"t value"[2], digits=2)`$.

***p*-value**

&emsp;&emsp; $`r p.value.string(one_t$"Pr(>|t|)"[2])`$.

**Rejection Region**

&emsp;&emsp; Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

&emsp;&emsp; Reject $H_0$. There is sufficient evidence to suggest that the regression line is significant.


### 95% confidence interval for $\beta_1$
```{r}
one_ci <- as_tibble(confint(one_model, level=0.95))
```

The confidence interval for the slope of age is (`r round(one_ci$"2.5 %"[2], digits = 2)`, `r round(one_ci$"97.5 %"[2], digits = 2)`).

### 95% confidence interval for $\mu_{Y|X_0}$ when $X_0=35$

```{r}

# want CI for mu when X_0=35, but this doesn't exist in dataset
age <- 35
sbp <- NA

new <- tibble(sbp, age)

# combine old and new data
one_cb <- rbind(one, new)

# find predictions and CI
one_cb$Prediction <- predict(one_model, newdata = one_cb)

one_cb$LCL <- predict(one_model, newdata = one_cb, 
                        interval = "confidence", 
                        level = 0.95)[, 2]
one_cb$UCL <- predict(one_model, newdata = one_cb, 
                        interval = "confidence", 
                        level = 0.95)[, 3]

# filter down to just X_0=35
one_ci_age35 <- filter(one_cb, age==35)

```
The confidence interval for $\mu_{Y|X_0}$ when $X_0=35$ is (`r round(one_ci_age35$LCL[1], digits = 2)`, `r round(one_ci_age35$UCL[1], digits = 2)`).


### 95% confidence band for $\mu_{Y|X_0}$

```{r}
p <- ggplot(one_cb, aes(x=age, y=Prediction)) +
  # Add a ribbon with the confidence band
  geom_smooth(color="black",
    aes(
      # lower and upper bound of the ribbon
      ymin = LCL, ymax = UCL
      ),
    stat = "identity") +
  xlab("Age") +
  ylab("SBP") +
  theme_bw()

p
```

### 95% prediction interval for $Y_{X_0}$ when $X_0=35$

```{r}

# want CI for mu when X_0=35, but this doesn't exist in dataset
age <- 35
sbp <- NA

new <- tibble(sbp, age)

# combine old and new data
one_pb <- rbind(one, new)

# find predictions and CI
one_pb$Prediction <- predict(one_model, newdata = one_cb)

one_pb$LCL <- predict(one_model, newdata = one_cb, 
                        interval = "prediction", 
                        level = 0.95)[, 2]
one_pb$UCL <- predict(one_model, newdata = one_cb, 
                        interval = "prediction", 
                        level = 0.95)[, 3]

# filter down to just X_0=35
one_pi_age35 <- filter(one_cb, age==35)

```
The prediction interval for $Y_{X_0}$ when $X_0=35$ is (`r round(one_pi_age35$LCL[1], digits = 2)`, `r round(one_pi_age35$UCL[1], digits = 2)`).


### 95% prediction band for $\mu_{Y|X_0}$

```{r}
p <- ggplot(one_pb, aes(x=age, y=Prediction)) +
  # Add a ribbon with the confidence band
  geom_smooth(color="black",
    aes(
      # lower and upper bound of the ribbon
      ymin = LCL, ymax = UCL
      ),
    stat = "identity") +
  xlab("Age") +
  ylab("SBP") +
  theme_bw()

p
```

### Scatterplot with regression line, 95% confidence bands, and 95% prediction bands

```{r}
# Redo predictions without X=35
one_pi <- predict(one_model, interval = "prediction")

# Combine dataset with PI 
new_one <- cbind(one, one_pi)

p <- ggplot(new_one, aes(x=age, y=sbp)) +
  geom_point() +
  stat_smooth(method = lm,color="black") +
  geom_line(aes(y = lwr), color = "red", linetype = "dashed") +
  geom_line(aes(y = upr), color = "red", linetype = "dashed") +
  xlab("Age") +
  ylab("SBP") +
  theme_bw()

p
```
